Progress Report

### 2023-09-18 Docs/ProgressReport.Rmd
explore glmmTMB

---
title: "Progress Report"
author: "Ying Jin"
date: "2023-05-16"
output: 
  html_document:
    self_contained: no
    number_sections: true
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
---

@@

- This is where near-unidentifiability issues first appear
- Though point estimation not necessarily affected

```{r, cache=TRUE, warning=FALSE}
# split by each bin
df <- df %>% rename(id = SEQN, Y=Z)
df_bin_lst <- split(df, f = df$bin)

# fit local GLMM and estimate latent function
# near-unidentifiability issues 
df_est_latent <- lapply(df_bin_lst, function(x){pred_latent(x)}) 
df_est_latent <- bind_rows(df_est_latent)

@@

```{r, message=FALSE}
df %>% 
  filter(id %in% rand_id) %>%
  left_join(df_est_latent, by = c("id", "bin")) %>%
  mutate(eta_hat = exp(eta_hat)/(1+exp(eta_hat))) %>%
  ggplot()+
  geom_line(aes(x=sind, y=eta_hat, group = id))+
  geom_point(aes(x=sind, y = Y, group = id), size = 0.5)+
  facet_wrap(~id, scales = "free")+
  geom_vline(xintercept = c(465, 545, 555, 1185, 1195, 1205), col = "red",
             alpha = 0.5, linewidth = 0.5)+
  labs(x = "Time", y = "Estimated latent function (probablity scale)")
```

```{r, message=FALSE}
# overview at estimated latent function
df %>% 
  filter(id %in% rand_id) %>%
  group_by(id, bin) %>%
  summarise(num = sum(Y)) %>%
  left_join(df_est_latent, by = c("id", "bin")) %>%
  # mutate(eta_hat = exp(eta_hat)/(1+exp(eta_hat))) %>%
  ggplot()+
  geom_line(aes(x=bin, y=eta_hat, group = id, linetype = "dashed"))+
  geom_point(aes(x=bin, y = num, group = id), size = 0.5)+
  facet_wrap(~id, scales = "free")+
  geom_vline(xintercept = c(465, 545, 555, 1185, 1195, 1205), col = "red",
             linewidth = 0.5, alpha = 0.6)+
  labs(x = "Bin", y = "Estimated latent function/number of active minutes")
```

- Take a closer look at the unidentifiable bins?

I just realized glmer actually used Laplace approximation as well. Could this also be due to numeric stability? In some bins (465, 545, 555, 1185, 1195, 1205), the approximation failed to converge.

Below is the warning messages and one model summary.

For this bin, the problem is with the fixed effect standard error. It is clearly near-zero.

https://www.learn-mlms.com/07-module-7.html

https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html

```{r, eval=FALSE}
try_fit_lst<-list()
for(i in seq_along(df_bin_lst)){
  this_glm <- tryCatch({glmer(Y ~ 1 + (1|id), data = df_bin_lst[[i]],
                             family = binomial)},
                       warning=function(w){NA})
  try_fit_lst[[i]] <- this_glm

@@

# bins where glmer returned errors
# names(df_bin_lst)[is.na(try_fit_lst)]
```

```{r}
issue_fit <- glmer(Y ~ 1 + (1|id), data = df_bin_lst[['465']],
                             family = binomial)
summary(issue_fit)

# table(df_bin_lst[['5']]$id, df_bin_lst[['5']]$Y)
```

```{r eval=FALSE}
colSums(table(df_bin_lst[['5']]$id, df_bin_lst[['5']]$Y)==0)
colSums(table(df_bin_lst[['465']]$id, df_bin_lst[['465']]$Y)==0)
```

```{r}
mat_est_unique <- matrix(df_est_latent$eta_hat, nrow=N, ncol=n_bin, byrow = F) 
# row index subject, column binned time
fpca_mod <- fpca.face(mat_est_unique, pve = 0.95, argvals = mid, var=T)

K <- ncol(fpca_mod$efunctions) # 18 eigenfunctions
```

```{r}
# population mean
plot(mid, fpca_mod$mu, type = "l", xlab = "bin", ylab = "mean")
```

```{r}
# eigenfunctions
par(mfrow=c(2,2))
plot(mid, fpca_mod$efunctions[, 1], type="l", xlab="bin", ylab="PC1")
plot(mid, fpca_mod$efunctions[, 2], type="l", xlab="bin", ylab="PC2")
plot(mid, fpca_mod$efunctions[, 3], type="l", xlab="bin", ylab="PC3")
plot(mid, fpca_mod$efunctions[, 4], type="l", xlab="bin", ylab="PC4")
```

```{r}
# covariance
heatmap(fpca_mod$VarMats[[which(unique(df_est_latent$id)==rand_id[1])]],
        Rowv = NA, Colv = NA, main = "Covariance matrix")
```

### Maximum likelihood

- Boundary hitting issues with numeric maximization
- Warning message: Steady solution not reached
- Large variation in both predicted values and scores. The numeric solution from maximum likelihood is obviously not stable. It is very likely to fail to converge.
- Perhaps it can be fixed with numeric options? But I'd rather not go that route because it depend on chance a lot.

```{r}
# Out-of-sample prediction
# maximum observations time: 360, 720, 1080
df_pred <- df_est_latent %>% filter(id %in% rand_id)
df_pred[, 'pred_t360'] <- df_pred[, 'pred_t720'] <- df_pred[, 'pred_t1080'] <- NA
score_out_mat <- array(NA, dim = c(length(rand_id), K, 3))
# dim indexes subject, eigenfunction and max obs time respectively
```

```{r, warning=FALSE}
# prediction for a single subject
for(i in rand_id){
  df_i <- df %>% filter(id==i) %>% mutate(bin=as.numeric(bin))
  # prediction
  pred_t360 <- out_samp_dyn_pred(df_new = df_i %>% filter(sind<=360),
                                 fpca_fit = fpca_mod, K = K)
  pred_t720 <- out_samp_dyn_pred(df_new = df_i %>% filter(sind<=720),
                                 fpca_fit = fpca_mod, K = K)
  pred_t1080 <- out_samp_dyn_pred(df_new = df_i %>% filter(sind<=1080),
                                  fpca_fit = fpca_mod, K = K)
  # prediction in container
  df_pred[df_pred$id==i, 'pred_t360'] <- pred_t360$eta_pred
  df_pred[df_pred$id==i, 'pred_t720'] <- pred_t720$eta_pred
  df_pred[df_pred$id==i, 'pred_t1080'] <- pred_t1080$eta_pred
  # score in container
  # score_out_mat[i, ,1] <- pred_t360$score_out
  # score_out_mat[i, ,2] <- pred_t720$score_out
  # score_out_mat[i, ,3] <- pred_t1440$score_out
}

```{r}
df_pred$pred_t360[df_pred$bin<=360] <- NA
df_pred$pred_t720[df_pred$bin<=720] <- NA
df_pred$pred_t1080[df_pred$bin<=1080] <- NA
# df_pred %>% filter(id == rand_id[1])
```

```{r, warning=FALSE}
df_pred %>%
  ggplot()+
  geom_line(aes(x=bin, y=eta_hat, group = id))+
  geom_line(aes(x=bin, y = pred_t360, col = "360", group = id))+
  geom_line(aes(x=bin, y = pred_t720, col = "720", group = id))+
  geom_line(aes(x=bin, y = pred_t1080, col = "1080", group = id))+
  facet_wrap(~id)
```

### Bayes method like Laplace approximation

```{r, warning=FALSE, message=FALSE, results='hide'}
df_pred2 <- df_est_latent %>% filter(id %in% rand_id)
df_pred2[, 'pred_t360'] <- df_pred2[, 'pred_t720'] <- df_pred2[, 'pred_t1080'] <- NA
score_out_mat2 <- array(NA, dim = c(length(rand_id), K, 3))

# unique id
# prediction for a single subject
source(here("Code/OutsampBayes.R"))

@@

  # prediction 
  ## up to 360
  pred_t1 <- out_pred_laplace(fpca_mod, df_i %>% filter(sind<=360))
  eta_pred_t1 <- fpca_mod$mu+fpca_mod$efunctions%*%pred_t1$score_out
  df_pred2[df_pred2$id == i, 'pred_t360'] <- eta_pred_t1
  ## up to 720
  pred_t2 <- out_pred_laplace(fpca_mod, df_i %>% filter(sind<=720))
  eta_pred_t2 <- fpca_mod$mu+fpca_mod$efunctions%*%pred_t2$score_out
  df_pred2[df_pred2$id == i, 'pred_t720'] <- eta_pred_t2
  ## up to 1080
  pred_t3 <- out_pred_laplace(fpca_mod, df_i %>% filter(sind<=1080))
  eta_pred_t3 <- fpca_mod$mu+fpca_mod$efunctions%*%pred_t3$score_out
  df_pred2[df_pred2$id == i, 'pred_t1080'] <- eta_pred_t3

}
```

```{r}
df_pred2$pred_t360[df_pred2$bin<=360] <- NA
df_pred2$pred_t720[df_pred2$bin<=720] <- NA
df_pred2$pred_t1080[df_pred2$bin<=1080] <- NA
# df_pred %>% filter(id == rand_id[1])
```

```{r, warning=FALSE}
df_pred2 %>%
  ggplot()+
  geom_line(aes(x=bin, y=eta_hat, group = id))+
  geom_line(aes(x=bin, y = pred_t360, col = "360", group = id))+
  geom_line(aes(x=bin, y = pred_t720, col = "720", group = id))+
  geom_line(aes(x=bin, y = pred_t1080, col = "1080", group = id))+
  facet_wrap(~id)
```

<!-- ## Failed prediction  -->

<!-- - Laplace approximation failed for some subjects -->
<!-- - Only when the observed track is short (up to 360) -->
<!-- - These subjects have something in common -->

<!-- ```{r} -->
<!-- # compare subjects that failed at Laplace Approximation with others -->
<!-- rand_id <- sample(unique(df$SEQN), size = 4) -->
<!-- com_id <- c(rand_id, sample(skip_id, 4)) -->

<!-- df_est_latent %>% filter(id %in% com_id) %>% -->
<!--   left_join(df %>% select(SEQN, Z, sind) %>% -->
<!--               rename(id = SEQN, Y=Z, bin = sind), by = c("id", "bin")) %>% -->
<!--   mutate(type = ifelse(id %in% rand_id, "Prediction", "Failed prediction")) %>% -->
<!--   ggplot(aes(x=bin, y=eta_hat, col = type))+ -->
<!--   geom_line()+ -->
<!--   geom_vline(xintercept = 360)+ -->
<!--   facet_wrap(~type+id, nrow = 2, ncol = 4)+ -->
<!--   labs(y = "Estimated latend function", x = "Time", col = "") -->
<!-- ``` -->

<!-- ## Subjects with complete prediction -->

<!-- ```{r fig_prob, warning=FALSE} -->
<!-- df_est_latent %>% -->
<!--   filter(id %in% rand_id) %>% -->
<!--   mutate(pred_t1080 = ifelse(bin<=1080, NA, pred_t1080), -->
<!--          pred_t720 = ifelse(bin<=720, NA, pred_t720), -->
<!--          pred_t360 = ifelse(bin<=360, NA, pred_t360)) %>% -->
<!--   mutate_at(vars(pred_t360, pred_t720, pred_t1080), function(x)exp(x)/(1+exp(x))) %>% -->
<!--   left_join(df %>% select(SEQN, Z, sind) %>% -->
<!--               rename(id = SEQN, Y=Z, bin = sind), by = c("id", "bin")) %>% -->
<!--   ggplot()+ -->
<!--     geom_line(aes(x=bin, y = pred_t360, col = "360"), linetype = "dashed")+ -->
<!--     geom_line(aes(x=bin, y = pred_t720, col = "720"),linetype = "dashed")+ -->
<!--     geom_line(aes(x=bin, y = pred_t1080, col = "1080"),linetype = "dashed")+ -->
<!--     geom_point(aes(x=bin, y = Y, col = "Outcome"), size = 0.5)+ -->
<!--     facet_wrap(~id)+ -->
<!--     labs(x = "Time", y = "Estimated latent function (probablity scale)", -->
<!--          col = "Maximum observation time") -->
<!-- ``` -->

<!-- ```{r fig_eta, warning=FALSE} -->
<!-- df_est_latent %>% -->
<!--   filter(id %in% rand_id) %>% -->
<!--   mutate(pred_t1080 = ifelse(bin<=1080, NA, pred_t1080), -->
<!--          pred_t720 = ifelse(bin<=720, NA, pred_t720), -->
<!--          pred_t360 = ifelse(bin<=360, NA, pred_t360)) %>% -->
<!--   ggplot()+ -->
<!--     geom_line(aes(x=bin, y = pred_t360, col = "360"), linetype = "dashed")+ -->
<!--     geom_line(aes(x=bin, y = pred_t720, col = "720"),linetype = "dashed")+ -->
<!--     geom_line(aes(x=bin, y = pred_t1080, col = "1080"),linetype = "dashed")+ -->
<!--     geom_line(aes(x=bin, y = eta_hat, col = "eta_hat"), linetype = "solid" )+ -->
<!--     facet_wrap(~id)+ -->
<!--     labs(x = "Time", y = "Estimated latent function") -->
<!-- ``` -->

<!-- ## AUC on subjects with complete prediction -->

<!-- - Use interpolation to fill in gaps between bins -->

<!-- ```{r} -->
<!-- # subjects with complete prediction -->
<!-- N <- length(unique(df$SEQN))-length(skip_id) -->
<!-- eval_id <- setdiff(unique(df$SEQN), skip_id) -->
<!-- J <- max(df$sind) -->

<!-- # containter: row index subject, column index time -->
<!-- eta_mat_t360 <- eta_mat_t720 <- eta_mat_t1080 <- matrix(NA, nrow = N, ncol = J) -->

<!-- # interpolation -->
<!-- for(i in seq_along(eval_id)){ -->
<!--    eta_mat_t360[i, ] <- approx(x=df_est_latent$bin[df_est_latent$id==eval_id[i]], -->
<!--                                y=df_est_latent$pred_t360[df_est_latent$id==eval_id[i]],  -->
<!--                                xout = 1:J)$y -->
<!--    eta_mat_t720[i, ] <- approx(x=df_est_latent$bin[df_est_latent$id==eval_id[i]], -->
<!--                                y=df_est_latent$pred_t720[df_est_latent$id==eval_id[i]],  -->
<!--                                xout = 1:J)$y -->
<!--    eta_mat_t1080[i, ] <- approx(x=df_est_latent$bin[df_est_latent$id==eval_id[i]], -->
<!--                                y=df_est_latent$pred_t1080[df_est_latent$id==eval_id[i]],  -->
<!--                                xout = 1:J)$y -->
<!--   } -->

<!-- ``` -->

<!-- ```{r} -->
<!-- # true outcome -->
<!-- Y_mat <- df$Z[df$SEQN %in% eval_id] -->
<!-- Y_mat <- matrix(Y_mat, nrow = N, ncol = J, byrow = T)  -->
<!-- ``` -->

<!-- ```{r AUC_fGFPCA} -->
<!-- # maximum observations time -->
<!-- t_vec <- c(360, 720, 1080) -->

<!-- # AUC container -->
<!-- auc_mat <- matrix(NA, 3, 3) -->
<!-- colnames(auc_mat) <- c("360", "720", "1080") -->
<!-- rownames(auc_mat) <- c("(360, 720]", "(720, 1080]", "(1080, 1440]") -->

<!-- # Calculate AUC -->

<!-- auc_mat[1, 1] <- performance(prediction(as.vector(eta_mat_t360[, 361:720]),  -->
<!--                                    as.vector(Y_mat[, 361:720])), -->
<!--                         "auc")@y.values[[1]] -->
<!-- auc_mat[2, 1] <- performance(prediction(as.vector(eta_mat_t360[, 721:1080]),  -->
<!--                                    as.vector(Y_mat[, 721:1080])), -->
<!--                         "auc")@y.values[[1]] -->
<!-- auc_mat[3, 1] <- performance(prediction(as.vector(eta_mat_t360[, 1081:1435]),  -->
<!--                                    as.vector(Y_mat[, 1081:1435])), -->
<!--                         "auc")@y.values[[1]] -->

<!-- auc_mat[2, 2] <- performance(prediction(as.vector(eta_mat_t720[, 721:1080]),  -->
<!--                                    as.vector(Y_mat[, 721:1080])), -->
<!--                         "auc")@y.values[[1]] -->
<!-- auc_mat[3, 2] <- performance(prediction(as.vector(eta_mat_t720[, 1081:1435]),  -->
<!--                                    as.vector(Y_mat[, 1081:1435])), -->
<!--                         "auc")@y.values[[1]] -->

<!-- auc_mat[3, 3] <- performance(prediction(as.vector(eta_mat_t1080[, 1081:1435]),  -->
<!--                                    as.vector(Y_mat[, 1081:1435])), -->
<!--                         "auc")@y.values[[1]] -->
<!-- auc_mat -->
<!-- ``` -->

<!-- ## Reference method -->

<!-- - GLMMadaptive? -->
<!-- - fcr on lowess smoothed track? -->

>>>>

---
title: "Progress Report"
author: "Ying Jin"
date: "2023-09-15"
output: 
  html_document:
    self_contained: no
    number_sections: true
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
    font: 12pt
---

@@

Following up on last report, when using glmer, the results from PRILS (nAGQ=0) seems more in line with experiences. As validation, let's fit the marginal model and look at the estimated population-average latent function, and compare it with the FPCA mean function. Specifically, at bin s the model would be $g(E(Y_i(s)))=f_0(s)$, with no subject-specific random effect.

Another validation I came up with is to look at the local intercepts. That is the $\beta_0(s)$ in local mixed models $g(E(Y_i(s)))=\beta_0(s)+b_i(x)$. This should correspond to the FPCA mean functions, as the population average level.

Since the individual random effects are centered around zero, I believe the marginal $f_0(s)$ should look very much like the conditional $\beta_0(x)$ if the model local fit well. The result obviously is in favor of PRILS.n *But what is the reason for AGQ to overshoot downward?*

```{r}
# split by each bin
df <- df %>% rename(id = SEQN, Y=Z)
df_bin_lst <- split(df, f = df$bin)
```

```{r cache=TRUE}
# local marginal GLM
local_marg <- function(df){
    this_glm <- glm(Y ~ 1, data = df, family = binomial)
    eta_hat <- predict(this_glm, type = "link")
    df$eta_hat <- eta_hat
    return(df)
}

df_est_local_marg <- lapply(df_bin_lst, function(x){local_marg(x)})

# local conditional GLMM
## PIRLS vs AGQ
df_est_cond_mean_pirls <- lapply(df_bin_lst, function(x){mean_latent(x, n_node = 0)}) 
df_est_cond_mean_agq <- lapply(df_bin_lst, function(x){mean_latent(x, n_node = 5)}) 
```

```{r, fig.width=12, fig.height=4}
ggarrange(
df_est_local_marg %>% bind_rows() %>%
  select(bin, eta_hat) %>% distinct(.) %>%
  ggplot(aes(x=bin, y=eta_hat))+
  geom_line()+
  labs(x="local GLM (marginal)", y = "latent mean")+
  ylim(-25, 1),

data.frame(bin=mid, eta_hat = unlist(df_est_cond_mean_pirls)) %>%
  ggplot(aes(x=bin, y=eta_hat))+
  geom_line()+
  labs(x="local GLMM (conditional) using PRILS", y = "latent mean")+
  ylim(-25, 1),

data.frame(bin=mid, eta_hat = unlist(df_est_cond_mean_agq)) %>%
  ggplot(aes(x=bin, y=eta_hat))+
  geom_line()+
  labs(x="local GLMM (conditional) using AGQ", y = "latent mean")+
  ylim(-25, 1),

nrow = 1, align = "h"
)
```

Let's for now trust the results from PIRLS more than AGQ, if we are using glmer.

```{r cache=TRUE}
# fit local GLMM and estimate latent function
# to avoid near-unidentifiability issues, using PIRLS maximization with nAGQ = 0

# fit local models
df_est_latent <- lapply(df_bin_lst, function(x){pred_latent(x, n_node = 0)}) 
df_est_latent <- bind_rows(df_est_latent)

@@

```{r, message=FALSE}
df %>% 
  filter(id %in% rand_id) %>%
  left_join(df_est_latent, by = c("id", "bin")) %>%
  mutate(eta_hat = exp(eta_hat)/(1+exp(eta_hat))) %>%
  ggplot()+
  geom_line(aes(x=sind, y=eta_hat, group = id))+
  geom_point(aes(x=sind, y = Y, group = id), size = 0.5)+
  facet_wrap(~id, scales = "free")+
  labs(x = "Time", y = "Estimated latent function (probablity scale)")
```

```{r, message=FALSE}
# overview at estimated latent function
df %>% 
  filter(id %in% rand_id) %>%
  group_by(id, bin) %>%
  summarise(num = sum(Y)) %>%
  left_join(df_est_latent, by = c("id", "bin")) %>%
  # mutate(eta_hat = exp(eta_hat)/(1+exp(eta_hat))) %>%
  ggplot()+
  geom_line(aes(x=bin, y=eta_hat, group = id))+
  geom_point(aes(x=bin, y = num, group = id), size = 0.5)+
  facet_wrap(~id, scales = "free")+
  labs(x = "Bin", y = "Estimated latent function/number of active minutes")
```

#### An alternative: glmmTMB

The default optimizer of glmmTMB is nlminb (quasi-Newton BFGS), whereas glmer uses bobyaqa (nAGQ=0) or Nelder_Mead (nAGQ>0)?

Let's first try to re-do the who local GLMM step using glmmTMB. Looks like no numeric issue occurred. It takes slightly longer time than PIRLS in glmer

```{r}
library(glmmTMB)
# re-fit local GLMM and check numeric issues
try_fit_lst<-list()
t1=Sys.time()
for(i in seq_along(df_bin_lst)){
  # this_glm <- tryCatch({glmer(Y ~ 1 + (1|id), data = df_bin_lst[[i]], family = binomial, nAGQ=5)},
  #                      warning=function(w){NA})
  this_glm <- glmmTMB(Y ~ 1 + (1|id), data = df_bin_lst[[i]], family = binomial)
  try_fit_lst[[i]] <- this_glm

}
df_est_latent_tmb <- lapply(try_fit_lst, function(x){predict(x, type = "link")})
t2=Sys.time()
t_tmb <- t2-t1
# population mean
```

- Time: `r t_tmb` minutes
- Mean/Intercept function: It looks like glmmTMB estimates lie right between PIRLS and AGQ! It takes slightly shorter time than PIRLS and mean function shoots more downward, not as bad as AGQ though.

```{r, fig.width=12, fig.height=4}
mean_tmb <- sapply(try_fit_lst, function(x){coef(summary((x)))$cond[1]})

ggarrange(
df_est_local_marg %>% bind_rows() %>%
  select(bin, eta_hat) %>% distinct(.) %>%
  ggplot(aes(x=bin, y=eta_hat))+
  geom_line()+
  labs(x="local GLM (marginal)", y = "latent mean")+
  ylim(-15, 1),

data.frame(bin=mid, eta_hat = unlist(df_est_cond_mean_pirls)) %>%
  ggplot(aes(x=bin, y=eta_hat))+
  geom_line()+
  labs(x="local GLMM (conditional) using PRILS (glmer)", y = "latent mean")+
  ylim(-15, 1),

data.frame(bin=mid, eta_hat = mean_tmb) %>%
  ggplot(aes(x=bin, y=eta_hat))+
  geom_line()+
  labs(x="local GLMM (conditional) using glmmTMB", y = "latent mean")+
  ylim(-15, 1),

nrow = 1
)
```

- Individual latent functions

```{r}
for(i in seq_along(df_est_latent_tmb)){
  df_est_latent_tmb[[i]] <- data.frame(df_bin_lst[[i]], eta_hat = df_est_latent_tmb[[i]])
}

df_est_latent_tmb <- bind_rows(df_est_latent_tmb)

# keep midpoint values
df_est_latent_tmb <- df_est_latent_tmb %>%
  select(-sind, -Y) %>% distinct(.)

```{r, message=FALSE, fig.height=8, fig.width=16}
df %>% 
  filter(id %in% rand_id) %>%
  left_join(df_est_latent %>% select(id, bin, eta_hat), by = c("id", "bin")) %>%
  left_join(df_est_latent_tmb %>% select(id, bin, eta_hat) %>% rename(eta_hat_tmb=eta_hat), 
            by = c("id", "bin")) %>%
  mutate_at(vars(eta_hat, eta_hat_tmb), function(x){exp(x)/(1+exp(x))}) %>% 
  ggplot()+
  geom_line(aes(x=sind, y=eta_hat, col = "glmer", alpha = 0.5))+
  geom_line(aes(x=sind, y=eta_hat_tmb, col = "glmmTMB", alpha = 0.5))+
  geom_point(aes(x=sind, y = Y), size = 0.5)+
  facet_wrap(~id, scales = "free")+
  labs(x = "Time", y = "Estimated latent function (probablity scale)")
```

```{r, message=FALSE, fig.height=8, fig.width=16}
# overview at estimated latent function
df %>% 
  filter(id %in% rand_id) %>%
  group_by(id, bin) %>%
  summarise(num = sum(Y)) %>%
  left_join(df_est_latent %>% select(id, bin, eta_hat), by = c("id", "bin")) %>%
  left_join(df_est_latent_tmb %>% select(id, bin, eta_hat) %>% rename(eta_hat_tmb=eta_hat), 
            by = c("id", "bin")) %>%
  ggplot()+
  geom_line(aes(x=bin, y=eta_hat, col = "glmer", alpha = 0.5))+
  geom_line(aes(x=bin, y=eta_hat_tmb, col = "glmmTMB", alpha = 0.5))+
  geom_point(aes(x=bin, y = num), size = 0.5)+
  facet_wrap(~id, scales = "free")+
  labs(x = "Time", y = "Estimated latent function")
```

So it looks like glmmTMB still overshoot downwards. Between the two methods, the mid-day part (active part for most people) seems to be very similar. Difference lies mostly at both ends, with consecutive inactive minutes.

```{r, eval=FALSE}
# numeric values of individual latent function
df %>% 
  filter(id %in% rand_id) %>%
  group_by(id, bin) %>%
  summarise(num = sum(Y)) %>%
  left_join(df_est_latent %>% select(id, bin, eta_hat), by = c("id", "bin")) %>%
  left_join(df_est_latent_tmb %>% select(id, bin, eta_hat) %>% rename(eta_hat_tmb=eta_hat), 
            by = c("id", "bin")) %>%
  head(20)
```

Follow up from late time, let's use PIRLS estimator and fix the number of PC functions instead of using all of them. We start with four PC functions.

I will use estimated latent functions from both glmer (nAGQ=0) and glmmTMB, and compare their results.

```{r}
# glmer nAGQ=0
mat_est_unique <- matrix(df_est_latent$eta_hat, nrow=N, ncol=n_bin, byrow = F) 
fpca_mod <- fpca.face(mat_est_unique, argvals = mid, var=T)

# lmmTMB
mat_est_unique2 <- matrix(df_est_latent_tmb$eta_hat, 
                          nrow=N, ncol=n_bin, byrow = F) 
fpca_mod2 <- fpca.face(mat_est_unique2, argvals = mid, var=T)

# ncol(fpca_mod$efunctions) # 27 eigenfunctions
K <- 4
```

- Population mean function

Hmm...FPCA seems to shoot the population mean downward in both methods, more so for glmmTMB. It looks like more rounds of model would make the population mean function more variable. Does that make sense?

```{r}
# population mean
par(mfrow=c(1, 2))
plot(mid, fpca_mod$mu, type = "l", xlab = "bin", ylab = "mean", main = "glmer",
     ylim = c(-12, 1))
plot(mid, fpca_mod2$mu, type = "l", xlab = "bin", ylab = "mean", main = "glmmTMB", ylim = c(-12, 1))
```
- PC functions

```{r, fig.height=16, fig.width=8}
# eigenfunctions
par(mfrow=c(4,2))

plot(mid, fpca_mod$efunctions[, 1], type="l", xlab="bin", ylab="PC1", main = "glmer", ylim = c(-0.21, 0.13))
plot(mid, fpca_mod2$efunctions[, 1], type="l", xlab="bin", ylab="PC1", main = "glmmTMB", ylim = c(-0.21, 0.13))

plot(mid, fpca_mod$efunctions[, 2], type="l", xlab="bin", ylab="PC2", ylim = c(-0.2, 0.14))
plot(mid, fpca_mod2$efunctions[, 2], type="l", xlab="bin", ylab="PC2",ylim = c(-0.2, 0.14))

plot(mid, fpca_mod$efunctions[, 3], type="l", xlab="bin", ylab="PC3", ylim = c(-0.2, 0.2))
plot(mid, fpca_mod2$efunctions[, 3], type="l", xlab="bin", ylab="PC3",ylim = c(-0.2, 0.2))

plot(mid, fpca_mod$efunctions[, 4], type="l", xlab="bin", ylab="PC4", ylim = c(-0.14, 0.23))
plot(mid, fpca_mod2$efunctions[, 4], type="l", xlab="bin", ylab="PC4",
     ylim = c(-0.14, 0.23))
```

- Individual covariance matrix

Wow! This covariacne matrix changed a lot from when using all of the PCs! It still looks cyclic, but now off-diagnoal region seems much more stable. Fluctuations mostly centered around diagnoal line.

```{r}
# covariance
heatmap(fpca_mod$VarMats[[which(unique(df_est_latent$id)==rand_id[1])]],
        Rowv = NA, Colv = NA, main = "glmer")

heatmap(fpca_mod2$VarMats[[which(unique(df_est_latent$id)==rand_id[1])]],
        Rowv = NA, Colv = NA, main = "glmmTMB")

Still from last time, we're gonna focus on Bayes methods now!

We are also going to extend the observed track, so that there are some variation in the observed data (instead of all zeros). I am going to start with 9am (540), 1pm (780),  5pm (1020)

Extending the track seems to solve the problem with failed Laplace Approximation. Not sure. Should try on the full sample.

```{r, warning=FALSE, message=FALSE}
df_pred <- df_est_latent %>% filter(id %in% rand_id)
df_pred[, 'pred_t540'] <- df_pred[, 'pred_t780'] <- df_pred[, 'pred_t1020'] <- NA
score_out_mat <- array(NA, dim = c(length(rand_id), K, 3))
```

```{r, results='hide'}
# unique id
# prediction for a single subject
source(here("Code/OutsampBayes.R"))

# PRILS
for(i in rand_id){
  df_i <- df %>% filter(id==i)

  # prediction 
  ## up to 360
  pred_t1 <- out_pred_laplace(fpca_mod, df_i %>% filter(sind<=540), kpc=K)
  eta_pred_t1 <- fpca_mod$mu+fpca_mod$efunctions[, 1:K]%*%pred_t1$score_out
  df_pred[df_pred$id == i, 'pred_t540'] <- eta_pred_t1
  ## up to 720
  pred_t2 <- out_pred_laplace(fpca_mod, df_i %>% filter(sind<=780), kpc=K)
  eta_pred_t2 <- fpca_mod$mu+fpca_mod$efunctions[, 1:K]%*%pred_t2$score_out
  df_pred[df_pred$id == i, 'pred_t780'] <- eta_pred_t2
  ## up to 1080
  pred_t3 <- out_pred_laplace(fpca_mod, df_i %>% filter(sind<=1020), kpc = K)
  eta_pred_t3 <- fpca_mod$mu+fpca_mod$efunctions[, 1:K]%*%pred_t3$score_out
  df_pred[df_pred$id == i, 'pred_t1020'] <- eta_pred_t3

@@

df_pred$pred_t540[df_pred$bin<=540] <- NA
df_pred$pred_t780[df_pred$bin<=780] <- NA
df_pred$pred_t1020[df_pred$bin<=1020] <- NA
```

```{r, results='hide'}
# glmmTMB
df_pred2 <- df_est_latent %>% filter(id %in% rand_id)
df_pred2[, 'pred_t540'] <- df_pred2[, 'pred_t780'] <- df_pred2[, 'pred_t1020'] <- NA

@@

  # prediction 
  ## up to 360
  pred_t1 <- out_pred_laplace(fpca_mod2, df_i %>% filter(sind<=540), kpc=K)
  eta_pred_t1 <- fpca_mod2$mu+fpca_mod2$efunctions[, 1:K]%*%pred_t1$score_out
  df_pred2[df_pred$id == i, 'pred_t540'] <- eta_pred_t1
  ## up to 720
  pred_t2 <- out_pred_laplace(fpca_mod2, df_i %>% filter(sind<=780), kpc=K)
  eta_pred_t2 <- fpca_mod2$mu+fpca_mod2$efunctions[, 1:K]%*%pred_t2$score_out
  df_pred2[df_pred$id == i, 'pred_t780'] <- eta_pred_t2
  ## up to 1080
  pred_t3 <- out_pred_laplace(fpca_mod2, df_i %>% filter(sind<=1020), kpc = K)
  eta_pred_t3 <- fpca_mod2$mu+fpca_mod2$efunctions[, 1:K]%*%pred_t3$score_out
  df_pred2[df_pred$id == i, 'pred_t1020'] <- eta_pred_t3

df_pred2$pred_t540[df_pred$bin<=540] <- NA
df_pred2$pred_t780[df_pred$bin<=780] <- NA
df_pred2$pred_t1020[df_pred$bin<=1020] <- NA
```

```{r, warning=FALSE, fig.width=16, fig.height=8}
ggarrange(
df_pred %>%
  ggplot()+
  geom_line(aes(x=bin, y=eta_hat, group = id))+
  geom_line(aes(x=bin, y = pred_t540, col = "540", group = id))+
  geom_line(aes(x=bin, y = pred_t780, col = "780", group = id))+
  geom_line(aes(x=bin, y = pred_t1020, col = "1020", group = id))+
  facet_wrap(~id,scales = "free")+labs(title = "glmer"),

df_pred2 %>%
  ggplot()+
  geom_line(aes(x=bin, y=eta_hat, group = id))+
  geom_line(aes(x=bin, y = pred_t540, col = "540", group = id))+
  geom_line(aes(x=bin, y = pred_t780, col = "780", group = id))+
  geom_line(aes(x=bin, y = pred_t1020, col = "1020", group = id))+
  facet_wrap(~id,scales = "free")+labs(title = "glmmTMB"),
nrow = 1, common.legend = T)
```
